#Here are my sample linea models. The dataframe needs to be changed on them to the regional datasets that mark created. i.e. using the "ranked" dataframe.



```{r corr, results='markup', eval=F}
library(Hmisc)

#Correlation Matrices per Year

#2016 Correlation Plot
dframe2016_1<-subset(dframe2016, select = c("mental_health_days", "mental_distress_rate", "inequality", "median_inc", "hs_grad", "college", "unempl", "child_poverty","single_parent", "severe_housing", "food_index","mh_providers","pop_provider_ratio"))

pairs(dframe2016_1)
a<-as.matrix(dframe2016_1)
b<-cor(a)
corrplot(b, is.corr=TRUE, method="number", title="2016 Correlation Matrix for All Numeric Vars.",mar=c(0,0,1,0))


#2017 Correlation Plot
dframe2017_1<-subset(dframe2017, select = c("mental_health_days", "mental_distress_rate", "inequality", "median_inc", "hs_grad", "college", "unempl", "child_poverty","single_parent", "severe_housing", "food_index","mh_providers","pop_provider_ratio"))

pairs(dframe2017_1)
a<-as.matrix(dframe2017_1)
b<-cor(a)
corrplot(b, is.corr=TRUE, method="number", title="2017 Correlation Matrix for All Numeric Vars.",mar=c(0,0,1,0))

#2018 Correlation Plot
dframe2018_1<-subset(dframe2018, select = c("mental_health_days", "mental_distress_rate", "inequality", "median_inc", "hs_grad", "college", "unempl", "child_poverty","single_parent", "severe_housing", "food_index","mh_providers","pop_provider_ratio"))

pairs(dframe2018_1)
a<-as.matrix(dframe2018_1)
b<-cor(a)
corrplot(b, is.corr=TRUE, method="number",title="2018 Correlation Matrix for All Numeric Vars.", mar=c(0,0,1,0))

#2019 Correlation Plot
dframe2019_1<-subset(dframe2019, select = c("mental_health_days", "mental_distress_rate", "inequality", "median_inc", "hs_grad", "college", "unempl", "child_poverty","single_parent", "severe_housing", "food_index","mh_providers","pop_provider_ratio"))

pairs(dframe2019_1)
a<-as.matrix(dframe2019_1)
b<-cor(a)
corrplot(b, is.corr=TRUE, method="number",title="2019 Correlation Matrix for All Numeric Vars.", mar=c(0,0,1,0))

#2020 Correlation Plot
dframe2020_1<-subset(dframe2020, select = c("mental_health_days", "mental_distress_rate", "inequality", "median_inc", "hs_grad", "college", "unempl", "child_poverty","single_parent", "severe_housing", "food_index","mh_providers","pop_provider_ratio"))

pairs(dframe2020_1)
a<-as.matrix(dframe2020_1)
b<-cor(a)
corrplot(b, is.corr=TRUE, method="number",title="2020 Correlation Matrix for All Numeric Vars.", mar=c(0,0,1,0))

#2021 Correlation Plot
dframe2020_1<-subset(dframe2021, select = c("mental_health_days", "mental_distress_rate", "inequality", "median_inc", "hs_grad", "college", "unempl", "child_poverty","single_parent", "severe_housing", "food_index","mh_providers","pop_provider_ratio"))

pairs(dframe2021_1)
a<-as.matrix(dframe2021_1)
b<-cor(a)
corrplot(b, is.corr=TRUE, method="number",title="2021 Correlation Matrix for All Numeric Vars.", mar=c(0,0,1,0))

```



```{r}
library(broom)

#Linear Models

#Linear Models for Mental Health Days (2016-2021)
lm_model_1 <- lm(mental_health_days~child_poverty, data=dframe)
summary(lm_model_1)

lm_model_2 <- lm(mental_health_days~child_poverty+median_inc+inequality+unempl+single_parent+severe_housing+food_index+mh_providers+pop_provider_ratio, data=dframe)
summary(lm_model_2)

#Linear Models for Mental Distress Rate (2016-2021)
lm_model_3 <- lm(mental_distress_rate~child_poverty, data=dframe)
summary(lm_model_3)

lm_model_4 <- lm(mental_distress_rate~child_poverty+median_inc+inequality+unempl+single_parent+severe_housing+food_index+mh_providers+pop_provider_ratio, data=dframe)
summary(lm_model_4)

#aov

#aov: Linear Models for Mental Health Days (2016-2021)
aov1 <- aov(lm_model_1)

table3<-(tidy(aov1))
table3 %>%
  kbl(caption="ANOVA Test for Model 1: lm(mental_health_days~child_poverty)",
       format= "html", col.names = c(" ","Df", "Sum Sq", "Mean Sq", "F Value", "Pr(>r)"), 
       align="r") %>% 
  kable_classic_2(full_width = F, html_font = "helvetica")


aov2 <- aov(lm_model_2)
table4<-(tidy(aov2))
table4 %>%
  kbl(caption="ANOVA Test for Model 2: lm(mental_health_days~child_poverty+median_inc+...+pop_provider_ratio)",
       format= "html", col.names = c(" ","Df", "Sum Sq", "Mean Sq", "F Value", "Pr(>r)"), 
       align="r") %>% 
  kable_classic_2(full_width = F, html_font = "helvetica")


#aov: Linear Models for Mental Health Days (2016-2021)
aov3 <- aov(lm_model_3)

table5<-(tidy(aov3))
table5 %>%
  kbl(caption="ANOVA Test for Model 3: lm(mental_distress_rate~child_poverty)",
       format= "html", col.names = c(" ","Df", "Sum Sq", "Mean Sq", "F Value", "Pr(>r)"), 
       align="r") %>% 
  kable_classic_2(full_width = F, html_font = "helvetica")


aov4 <- aov(lm_model_4)

table6<-(tidy(aov4))
table6 %>%
  kbl(caption="ANOVA Test for Model 4: lm(mental_distress_rate~child_poverty+median_inc+...+pop_provider_ratio)",
       format= "html", col.names = c(" ","Df", "Sum Sq", "Mean Sq", "F Value", "Pr(>r)"), 
       align="r") %>% 
  kable_classic_2(full_width = F, html_font = "helvetica")

