---
title: "Investigating the Relationship between Mental Health and Income Inequality"
author: "Team 6"
#date: "3/23/2022"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    number_sections: false
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
  runtime: shiny
---

```{r init, include=F}

# Import libraries
library(ezids)
library(dplyr)
library(data.table)
library(readr)
library(lattice)
library(corrplot)
library(psych)
library(kableExtra)
library(ggpubr)
library(Hmisc)
```


```{r setup, include=FALSE}

# knitr::opts_chunk$set(warning = F, results = "markup", message = F)
knitr::opts_chunk$set(warning = F, results = "hide", message = F, echo = TRUE)
options(scientific=T, digits = 3) 
# options(scipen=9, digits = 3) 
# ‘scipen’: integer. A penalty to be applied when deciding to print numeric values in fixed or exponential notation.  Positive values bias towards fixed and negative towards scientific notation: fixed notation will be preferred unless it is more than ‘scipen’ digits wider.
# use scipen=999 to prevent scientific notation at all times


```


# Technical Analysis

## Instructions

Technical analysis (R code), 25% (First attempt due: Mar 28 Monday 12:30pm)

Just before your team presentation, provide an R-markdown, or R script file, which shows the R-code and brief explanations as well as the rationale of the Exploratory Data Analysis of
your project. (Also submit your data file, or give the online source url.) This submission,
together with your final document (in part VI), will determine your team’s grade. This
document shows a technical person the math/stat/codes that you used in your analysis.

The document should include:

- Summary of the dataset
- Descriptive Statistics
- Graphical representations of the data
- [When applicable] Measures of Variance / sd
- [When applicable] Normality tests
- [When applicable] Initial correlation / Chi Square tests / ANOVA analysis / Z-test or Z-interval / T-test or T-interval etc.

## Summary

**Assigned to:** Mark

The variables in the dataset are:

* `region`: name of the US Census Bureau region (name)
* `division`: name of the US Census Bureau division (contained with a census region)
* `state`: state abbreviation
* `statecode`: FIPS state code
* `countycode`: FIPS county code
* `fipscode`: 5-digit FIPS Code (county-level); combines `statecode` and `countycode`
* `county`: county name
* `year`: report release year from [County Health Rankings](https://www.countyhealthrankings.org/); range of 2016-2021
* `county_ranked`: Indicates whether or not the county was ranked; 0=unranked, 1=ranked, or NA for aggregated national or state-level data
* `mental_health_days`: Average number of mentally unhealthy days reported in past 30 days (age-adjusted)
* `mental_distress_rate`: Percentage of adults reporting 14 or more days of poor mental health per month
* `inequality`: Ratio of household income at the 80th percentile to income at the 20th percentile (Income inequality)
* `median_inc`: The income where half of households in a county earn more and half of households earn less
* `hs_grad`: Percentage of adults ages 25 and over with a high school diploma or equivalent
* `college`: Percentage of adults ages 25-44 with some post-secondary education
* `unempl`: Percentage of population ages 16 and older unemployed but seeking work
* `child_poverty`: Percentage of people under age 18 in poverty
* `single_parent`: Percentage of children that live in a household headed by single parent
* `severe_housing`: Percentage of households with severe housing problems
* `food_index`: Index of factors that contribute to a healthy food environment, from 0 (worst) to 10 (best)
* `mh_providers`: rate of providers to 100,000 population
* `pop_provider_ratio`: ratio of population to mental health providers (i.e., population served per provider)
* `pop`: census population estimate
* `pct_below18`: percent of population younger than 18
* `pct_black`: percent of population that are African-American or non-Hispanic Black
* `pct_native_am`: percent of population that are Native American or Alaska Natives
* `pct_asian`: percent of population that are Asian
* `pct_pacific`: percent of population that are Native Hawaiian or Other Pacific Islander
* `pct_hispanic`: percent of population that are Hispanic
* `pct_white`: percent of population that are non-Hispanic white or Caucasian
* `pct_female`: percent of population that are female
* `pct_rural`: percent of population that live in rural areas

For more information, see [the measures online](https://www.countyhealthrankings.org/explore-health-rankings/measures-data-sources/2021-measures).


```{r import, results='hide', echo=F}

# import csv as dataframe
dframe1 <- data.frame(read.csv("data/processed/analytic_data_2016_2021.csv"))
View(dframe1)

dframe <- data.frame(read.csv("data/processed/analytic_data_2016_2021_with_regions.csv"))
View(dframe)

# convert variables to factor
dframe$region <- factor(dframe$region)
dframe$division <- factor(dframe$division)
dframe$statecode <- factor(dframe$statecode)
dframe$countycode <- factor(dframe$countycode)
dframe$fipscode <- factor(dframe$fipscode)
dframe$county_ranked <- factor(dframe$county_ranked)

# check structure of data
str(dframe)
summary(dframe)

```

```{r summary, results='hide'}

# look at county_ranked var; not all counties are ranked; also some aggregated data per state and country exist in the observations
# =1 means they are ranked, =0 means unranked, and =NA is for state/national data
print(summary(dframe$county_ranked))

# subset of dataframe including only ranked counties
ranked <- dframe %>% subset(county_ranked==1)

# subset of dataframe including only ranked counties
unranked <- dframe %>% subset(county_ranked==0)

# subset of dataframe including only aggregated data
aggregated <- dframe %>% subset(is.na(county_ranked))

# subset ranked data by region
ranked_MW <- ranked %>% subset(region=="Midwest")
ranked_NE <- ranked %>% subset(region=="Northeast")
ranked_SO <- ranked %>% subset(region=="South")
ranked_WE <- ranked %>% subset(region=="West")

# subset ranked data into annual datasets
ranked16 <- ranked %>% subset(year==2016)
ranked17 <- ranked %>% subset(year==2017)
ranked18 <- ranked %>% subset(year==2018)
ranked19 <- ranked %>% subset(year==2019)
ranked20 <- ranked %>% subset(year==2020)
ranked21 <- ranked %>% subset(year==2021)

# view size of annual datasets
df_annual_list <- list(ranked16, ranked17, ranked18, ranked19, ranked20, ranked21)
for (df in df_annual_list) {
  print(paste("Observations in", median(df$year), "data: ", nrow(df)))
  
}

# view head and tail of ranked data
xkabledplyhead(ranked, 2, title = "Table: First 2 Rows of Ranked Data")
xkabledplytail(ranked, 2, title = "Table: Last 2 Rows of Ranked Data")

```

Our data are identified at the county-year level. Our data set contains `r nrow(dframe)` observations and `r ncol(dframe)` variables, although `r nrow(aggregated)` of these observations are for aggregated data at the national or state level. In addition, `r nrow(unranked)` observations are for counties that are unranked by the [University of Wisconsin Population Health Institute](https://www.countyhealthrankings.org/explore-health-rankings/faq-page), suggesting that the data for these counties is less reliable.

In total, we have `r nrow(ranked)` observations in the ranked data, combined across 6 annual reports (2016–2021). Each annual data set has between 3071 and 3084 observations, indicating that the number of ranked counties is consistent over time.

We also grouped the data on several dimensions using the `group_by()` and `summarise()` functions from the dplyr package. This helped us better investigate time and geographic patterns in our data.

First, we grouped the data set by report year and calculating the population-weighted means for several variables. In general, the mean income inequality has stayed relatively constant over time, while the mental health variables have risen slightly. Other economic variables, such as unemployment and median income, have moved in positive directions (unemployment has fallen, while median incomes have increased). 

Next, we grouped the data by year and state and calculated population-weighted means and state-level medians. There was a lot more variation among the data at this level, which will be helpful when plotting aggregated data later.

Finally, we grouped the data by region, giving us a sense of the size of the 4 census regions: the South contains the most states and counties and has the largest population, while the Northeast is the smallest on these measures. The Northeast has highest measured inequality, and the Midwest has the lowest. The mental health variables are highest (i.e., worst) in the South and lowest in the West. These general trends hold up when grouping by year and region. These differences suggest that comparing the relationship between the variables across regions would be worthwhile.

```{r ranked, results='markup', message=FALSE}

# groupby year
ranked_by_year <- ranked %>% group_by(year)

summ_by_year <- ranked_by_year %>%  summarize(num_counties = n_distinct(fipscode), 
                              num_states = n_distinct(statecode),
                              wmean_inequality = weighted.mean(inequality, pop, na.rm=T), 
                              wmean_mh_rate = weighted.mean(mental_distress_rate, pop, na.rm=T), 
                              wmean_mh_days = weighted.mean(mental_health_days, pop, na.rm=T), 
                              wmean_unempl = weighted.mean(unempl, pop, na.rm=T), 
                              wmean_medinc = weighted.mean(median_inc, pop, na.rm=T),
                              )

xkabledplyhead(summ_by_year, 6, title = "Table: Data grouped by Year")


# groupby year and state
ranked_by_year_state <- ranked %>% group_by(year, statecode)

summ_by_year_state <- ranked_by_year_state %>% summarize(num_obs = n(), 
                                   state_name = first(state),
                                   region = first(region),
                                   wmean_inequality = weighted.mean(inequality, 
                                                                    pop, na.rm = T),
                                   med_inequality = median(inequality, na.rm = T),
                                   wmean_mh_rate = weighted.mean(mental_distress_rate, 
                                                                 pop, na.rm=T),
                                   med_mh_rate = median(mental_distress_rate, na.rm=T),
                                   wmean_mh_days = weighted.mean(mental_health_days, 
                                                                 pop, na.rm=T), 
                                   med_mh_days = median(mental_health_days, na.rm=T),
                                   wmean_unempl = weighted.mean(unempl, 
                                                                pop, na.rm=T), 
                                   med_unempl = median(unempl, na.rm=T),
                                   wmean_medinc = weighted.mean(median_inc, 
                                                                pop, na.rm=T),
                                   med_medinc = median(median_inc, na.rm=T)
                                   )

xkabledplyhead(summ_by_year_state, 8, title = "Table: Data grouped by Year and State")


# groupby region
ranked_by_region <- ranked %>% group_by(region)

summ_by_region <- ranked_by_region %>%  summarize(num_counties = n_distinct(fipscode), 
                              num_states = n_distinct(statecode),
                              wmean_inequality = weighted.mean(inequality, pop, na.rm=T), 
                              wmean_mh_rate = weighted.mean(mental_distress_rate, pop, na.rm=T), 
                              wmean_mh_days = weighted.mean(mental_health_days, pop, na.rm=T), 
                              wmean_unempl = weighted.mean(unempl, pop, na.rm=T), 
                              wmean_medinc = weighted.mean(median_inc, pop, na.rm=T),
                              )

xkabledplyhead(summ_by_region, 4, title = "Table: Data grouped by Region")


# groupby year and state
ranked_by_year_region <- ranked %>% group_by(year, region)

summ_by_year_region <- ranked_by_year_region %>% summarize(num_obs = n(), 
                                   region_name = first(region),
                                   sum_pop_millions = sum(pop/1000000, na.rm = T),
                                   wmean_inequality = weighted.mean(inequality, pop, na.rm = T),
                                   wmean_mh_rate = weighted.mean(mental_distress_rate, pop, na.rm=T),
                                   wmean_mh_days = weighted.mean(mental_health_days, pop, na.rm=T)
                                   )

xkabledplyhead(summ_by_year_region, 8, title = "Table: Data grouped by Year and Region")

```


## Descriptive Statistics

**Assigned to:** Alex

```{r dstats, results='markup'}
#Summary Statistics of all Variables used in dframe dataset
table1<-describeBy(dframe)
table1 %>%
  kbl(caption="Table 1: Summary Statistics for Master Dataset",
       format= "html", col.names = c("Var Num.","Count","Mean","Std. Dev.","Median", "Trimmed Mean", "Mad", "Minimum", "Maximum","Range", "Skewness", "Kurtosis", "S.E."),
      align="r") %>%
   kable_minimal(full_width = F, html_font = "helvetica")
#Summary Statistics of relevant Variables used in analysis below in dfram dataset
dframe2<-subset(dframe, select = c("mental_health_days", "mental_distress_rate", "inequality", "median_inc", "hs_grad", "college", "unempl", "child_poverty","single_parent", "severe_housing", "food_index","mh_providers","pop_provider_ratio","pct_rural"))
table2<-describeBy(dframe2)
table2 %>%
  kbl(caption="Table 1: Summary Statistics for Master Dataset",
       format= "html", col.names = c("Var Num.","Count","Mean","Std. Dev.","Median", "Trimmed Mean", "Mad", "Minimum", "Maximum","Range", "Skewness", "Kurtosis", "S.E."),
      align="r") %>%
   kable_minimal(full_width = F, html_font = "helvetica")
```


## Graphs and Figures

**Assigned to:** Mark (with help from Shumel & Xuan)

For example, histograms, boxplots, scatterplots, line graphs (for aggregated annual data), etc. We should use the `ggplot()` function.

```{r graphs_init, results='hide'}

loadPkg("ggplot2")
loadPkg("patchwork") ## for combining graphs
loadPkg("purrr")
loadPkg("tidyr")
loadPkg("gridExtra")

```

To visualize trends over time, we graphed several variables from the data, grouped by report year. Both measures of mental health have increased, but it's difficult to tell whether these increases are substantial. Notably, income inequality appears largely flat, while median income and unemployment have slightly improved.

```{r lines_annual, results='markup', message=F}

# use summ_by_year

# annual trends for mental health var
p1 <- ggplot(summ_by_year, aes(x=year)) +
  geom_line(aes(y = wmean_mh_days), lwd=1) + 
  expand_limits(y=c(0,5)) +
  labs(title = "Poor Mental Health Days", 
       x = "Report Year", y = "Avg for past 30 days") +
  theme_minimal()

p2 <- ggplot(summ_by_year, aes(x=year)) +
  geom_line(aes(y = wmean_mh_rate), lwd=1) + 
  expand_limits(y=c(0,.2)) +
  labs(title = "Frequent Mental Distress Rate", 
       x = "Report Year", y = "Percentage of Adults") + 
  theme_minimal()

# combine plots using patchwork package
p1 + p2

# annual trends for inequality and median income
p3 <- ggplot(summ_by_year, aes(x=year)) +
  geom_line(aes(y = wmean_inequality), lwd=1) + 
  expand_limits(y=c(0,5)) +
  labs(title = "Income Inequality", 
       x = "Report Year", y = "Ratio 80th : 20th pctl") +
  theme_minimal()

p4 <- ggplot(summ_by_year, aes(x=year)) +
  geom_line(aes(y = wmean_medinc / 1000), lwd=1) + 
  expand_limits(y=c(0,max(summ_by_year$wmean_medinc/1000)+5)) +
  labs(title = "Median Income", 
       x = "Report Year", y = "Thousands of Dollars") +
  theme_minimal()

p5 <- ggplot(summ_by_year, aes(x=year)) +
  geom_line(aes(y = wmean_unempl), lwd=1) + 
  expand_limits(y=c(0,.10)) +
  labs(title = "Unemployment Rate", 
       x = "Report Year", y = "Percent Unemployed") +
  theme_minimal()

# combine plots
p3 / (p4 + p5)

grid.arrange(p1,p2, nrow=1, ncol=2)
grid.arrange(p3,p4,p5, nrow=1, ncol=3)

```

Next, we graphed the county-level data using scatterplots to compare the mental health measures with economic variables.

Both poor mental health days and the frequent mental distress rate are positively correlated with income inequality. When coloring the scatterplot by median income, counties with a lower median income tend toward the outer limits of the figure, which suggests that poorer counties generally have higher inequality and worse mental health. It's harder to infer a clear relationship when coloring the points by `pct_rural`, although the many of the highly unequal counties are heavily urban.

There is not much of a correlation between the mental health outcomes and the ratio of the population to mental health providers.

```{r scatter_a, results='markup'}

# data: use ranked
# colnames(ranked)

#-- Dependent vs Explanatory --#

# color = median income
color_value <- "Median Inc. ($T)"

c1 <- ggplot(ranked, aes(y=mental_health_days, x=inequality, color=median_inc/1000)) + 
  geom_point() +
  geom_smooth(formula = y ~ x, method=lm, se=FALSE, color="#a89968") +  ## Add linear regression line
  labs(title = "Scatterplot of Poor Mental Health Days vs. Inequality",
       y = 'Poor Mental Health Days per 30', x = 'Income Inequality Rate', 
       color = color_value)

c2 <- ggplot(ranked, aes(y=mental_distress_rate, x=inequality, color=median_inc/1000)) + 
  geom_point() +
  geom_smooth(formula = y ~ x, method=lm, se=FALSE, color="#a89968") + 
  labs(title = "Scatterplot of Frequent Distress vs. Inequality",
       y = 'Mental Distress Rate', x = 'Income Inequality Rate', 
       color = color_value)

# color = % rural
color_value <- "% Rural Pop."

c3 <- ggplot(ranked, aes(y=mental_health_days, x=inequality, color=pct_rural)) + 
  geom_point() +
  geom_smooth(formula = y ~ x, method=lm, se=FALSE, color="#a89968") +  ## Add linear regression line
  labs(title = "Scatterplot of Poor Mental Health Days vs. Inequality",
       y = 'Poor Mental Health Days per 30', x = 'Income Inequality Rate', 
       color = color_value)

c4 <- ggplot(ranked, aes(y=mental_distress_rate, x=inequality, color=pct_rural)) + 
  geom_point() +
  geom_smooth(formula = y ~ x, method=lm, se=FALSE, color="#a89968") + 
  labs(title = "Scatterplot of Frequent Distress vs. Inequality",
       y = 'Mental Distress Rate', x = 'Income Inequality Rate', 
       color = color_value)


#-- Dependent vs. Provision of Mental Health Care --#

# color = % rural
color_value <- "% Rural Pop."

c5 <- ggplot(ranked, aes(y=mental_health_days, x=pop_provider_ratio, color=pct_rural)) + 
  geom_point() +
  geom_smooth(formula = y ~ x, method=lm, se=FALSE, color="#a89968") + 
  labs(title = "Scatterplot of Poor Mental Health Days vs. Mental Health Providers",
       y = 'Mental Distress Rate', x = 'Ratio Population : Providers', 
       color = color_value)

c6 <- ggplot(ranked, aes(y=mental_distress_rate, x=pop_provider_ratio, color=pct_rural)) + 
  geom_point() +
  geom_smooth(formula = y ~ x, method=lm, se=FALSE, color="#a89968") + 
  labs(title = "Scatterplot of Frequent Distress vs. Mental Health Providers",
       y = 'Mental Distress Rate', x = 'Ratio Population : Providers', 
       color = color_value)


grid.arrange(c1,c2, nrow=1, ncol=2)
grid.arrange(c3,c4, nrow=1, ncol=2)
grid.arrange(c5,c6, nrow=1, ncol=2)

```

To get a better sense of how income inequality is associated with the other economic variables, we created a figure with multiple scatterplots. In general, the variables suggest the expected relationship: for example, `inequality` is positively correlated with unemployment, severe housing issues, and child poverty but negatively associated with a better food environment and high school graduation rates. Both `mental_health_days` and `mental_distress_rate` exhibit similarly signed correlations with the economic variables.

```{r scatter_b, results='markup'}

# Multi-plot figures

#-- Inequality vs. other economic variables --#

# unempl, child_poverty, hs_grad, college, single_parent, severe_housing, food_index

gg1 <- ggplot(ranked, aes(y=inequality, x=unempl)) + 
  geom_point() + 
  geom_smooth(formula = y ~ x, method=lm, se=FALSE) + 
  labs(title = "Plot (a)")

gg2 <- ggplot(ranked, aes(y=inequality, x=severe_housing)) + 
  geom_point() +
  geom_smooth(formula = y ~ x, method=lm, se=FALSE) + 
  labs(title = "Plot (b)")

gg3 <- ggplot(ranked, aes(y=inequality, x=child_poverty)) + 
  geom_point() +
  geom_smooth(formula = y ~ x, method=lm, se=FALSE) + 
  labs(title = "Plot (c)")

gg4 <- ggplot(ranked, aes(y=inequality, x=single_parent)) + 
  geom_point() +
  geom_smooth(formula = y ~ x, method=lm, se=FALSE) + 
  labs(title = "Plot (d)")

gg5 <- ggplot(ranked, aes(y=inequality, x=hs_grad)) + 
  geom_point() +
  geom_smooth(formula = y ~ x, method=lm, se=FALSE) + 
  labs(title = "Plot (e)")

gg6 <- ggplot(ranked, aes(y=inequality, x=college)) + 
  geom_point() +
  geom_smooth(formula = y ~ x, method=lm, se=FALSE) + 
  labs(title = "Plot (f)")

gg7 <- ggplot(ranked, aes(y=inequality, x=food_index)) + 
  geom_point() +
  geom_smooth(formula = y ~ x, method=lm, se=FALSE) + 
  labs(title = "Plot (g)")

plot_list <- list(gg1, gg2, gg3, gg4, gg5, gg6, gg7)

do.call("grid.arrange", c(plot_list, ncol = 3))


#-- Dependent vs. other economic variables --#

# poor mental health days
gg1 <- ggplot(ranked, aes(y=mental_health_days, x=unempl)) + 
  geom_point() + 
  geom_smooth(formula = y ~ x, method=lm, se=FALSE) + 
  labs(title = "Plot (a)")

gg2 <- ggplot(ranked, aes(y=mental_health_days, x=severe_housing)) + 
  geom_point() +
  geom_smooth(formula = y ~ x, method=lm, se=FALSE) + 
  labs(title = "Plot (b)")

gg3 <- ggplot(ranked, aes(y=mental_health_days, x=child_poverty)) + 
  geom_point() +
  geom_smooth(formula = y ~ x, method=lm, se=FALSE) + 
  labs(title = "Plot (c)")

gg4 <- ggplot(ranked, aes(y=mental_health_days, x=single_parent)) + 
  geom_point() +
  geom_smooth(formula = y ~ x, method=lm, se=FALSE) + 
  labs(title = "Plot (d)")

gg5 <- ggplot(ranked, aes(y=mental_health_days, x=hs_grad)) + 
  geom_point() +
  geom_smooth(formula = y ~ x, method=lm, se=FALSE) + 
  labs(title = "Plot (e)")

gg6 <- ggplot(ranked, aes(y=mental_health_days, x=college)) + 
  geom_point() +
  geom_smooth(formula = y ~ x, method=lm, se=FALSE) + 
  labs(title = "Plot (f)")

gg7 <- ggplot(ranked, aes(y=mental_health_days, x=food_index)) + 
  geom_point() +
  geom_smooth(formula = y ~ x, method=lm, se=FALSE) + 
  labs(title = "Plot (g)")

plot_list <- list(gg1, gg2, gg3, gg4, gg5, gg6, gg7)

do.call("grid.arrange", c(plot_list, ncol = 3))


# mental distress rate
gg1 <- ggplot(ranked, aes(y=mental_distress_rate, x=unempl)) + 
  geom_point() + 
  geom_smooth(formula = y ~ x, method=lm, se=FALSE) + 
  labs(title = "Plot (a)")

gg2 <- ggplot(ranked, aes(y=mental_distress_rate, x=severe_housing)) + 
  geom_point() +
  geom_smooth(formula = y ~ x, method=lm, se=FALSE) + 
  labs(title = "Plot (b)")

gg3 <- ggplot(ranked, aes(y=mental_distress_rate, x=child_poverty)) + 
  geom_point() +
  geom_smooth(formula = y ~ x, method=lm, se=FALSE) + 
  labs(title = "Plot (c)")

gg4 <- ggplot(ranked, aes(y=mental_distress_rate, x=single_parent)) + 
  geom_point() +
  geom_smooth(formula = y ~ x, method=lm, se=FALSE) + 
  labs(title = "Plot (d)")

gg5 <- ggplot(ranked, aes(y=mental_distress_rate, x=hs_grad)) + 
  geom_point() +
  geom_smooth(formula = y ~ x, method=lm, se=FALSE) + 
  labs(title = "Plot (e)")

gg6 <- ggplot(ranked, aes(y=mental_distress_rate, x=college)) + 
  geom_point() +
  geom_smooth(formula = y ~ x, method=lm, se=FALSE) + 
  labs(title = "Plot (f)")

gg7 <- ggplot(ranked, aes(y=mental_distress_rate, x=food_index)) + 
  geom_point() +
  geom_smooth(formula = y ~ x, method=lm, se=FALSE) + 
  labs(title = "Plot (g)")

plot_list <- list(gg1, gg2, gg3, gg4, gg5, gg6, gg7)

do.call("grid.arrange", c(plot_list, ncol = 3))


```

Because we are interested in whether the relationship between mental health and income inequality differs geographically, we also colored the points in the scatterplot by region.

An initial look by region is consistent with our previous takeaways from the aggregated region data. In general, counties in the South seem to have both relatively poor mental health and greater income inequality, while counties in the Midwest seem to do best on both dimensions. 

To make sure this relationship is not sensitive to report year, we also graphed the data for 2016 and 2021 separately. Overall, the trends are very similar to the multi-year data. The 2016 data for mental_health_days does look strange though, almost as if it's rounded to one decimal place. To be safe, we plotted the 2017 report data, which also corresponded nicely to the other years.

```{r scatter_c, results='markup'}

# data: use ranked
color_value <- "Region"

#-- All Years --#

p1 <- ggplot(ranked, aes(y=mental_health_days, x=inequality, color = factor(region))) + 
  geom_point(show.legend = F) +
  labs(title = paste("(a) Per 30 days"),
       y = 'Poor Mental Health days per 30', x = 'Income Inequality Rate', 
       color = color_value)

p2 <- ggplot(ranked, aes(y=mental_distress_rate, x=inequality, color = factor(region))) + 
  geom_point() + 
  labs(title = paste("(b) Percentage"),
       y = 'Frequent Distress Rate', x = 'Income Inequality Rate', 
       color = color_value)

p_all <- p1 + p2 + plot_annotation(title = "Scatterplot of Mental Health Variables vs. Income Inequality, 2016-2021")
p_all


#-- Is relationship sensitive to report year? --#

# 2016 vs 2021

# 2016
p1 <- ggplot(ranked16, aes(y=mental_health_days, x=inequality, color = factor(region))) + 
  geom_point(show.legend = F) +
  labs(title = "(a) Per 30 days",
       y = 'Poor Mental Health Days per 30', x = 'Income Inequality Rate', 
       color = color_value)

p2 <- ggplot(ranked16, aes(y=mental_distress_rate, x=inequality, color = factor(region))) + 
  geom_point() + 
  labs(title = "(b) Percentage",
       y = 'Frequent Distress Rate', x = 'Income Inequality Rate', 
       color = color_value)

p16 <- p1 + p2 + plot_annotation(title = "Scatterplot of Mental Health Variables vs. Income Inequality, 2016")
p16

# 2021
p1 <- ggplot(ranked21, aes(y=mental_health_days, x=inequality, color = factor(region))) + 
  geom_point(show.legend = F) +
  labs(title = "(a) Per 30 days",
       y = 'Poor Mental Health Days per 30', x = 'Income Inequality Rate', 
       color = color_value)

p2 <- ggplot(ranked21, aes(y=mental_distress_rate, x=inequality, color = factor(region))) + 
  geom_point() + 
  labs(title = "(b) Percentage",
       y = 'Frequent Distress Rate', x = 'Income Inequality Rate', 
       color = color_value)

p21 <- p1 + p2 + plot_annotation(title = "Scatterplot of Mental Health Variables vs. Income Inequality, 2021")
p21


# 2016 mental_health_days data looks weird (like it's been rounded)
# let's check 2017
p1 <- ggplot(ranked17, aes(y=mental_health_days, x=inequality, color = factor(region))) + 
  geom_point(show.legend = F) +
  labs(title = "(a) Per 30 days",
       y = 'Poor Mental Health Days per 30', x = 'Income Inequality Rate', 
       color = color_value)

p2 <- ggplot(ranked17, aes(y=mental_distress_rate, x=inequality, color = factor(region))) + 
  geom_point() + 
  labs(title = "(b) Percentage",
       y = 'Frequent Distress Rate', x = 'Income Inequality Rate', 
       color = color_value)

p17 <- p1 + p2 + plot_annotation(title = "Scatterplot of Mental Health Variables vs. Income Inequality, 2017")
p17


```

The county-level scatterplots look crowded because of all the data. So we also plotted the aggregated data by state, using the population-weighted means and statewide medians. The relationship between mental health and income inequality at this aggregated level is consistent with the disaggregated scatterplots.

```{r scatter_agg, results='markup'}

# data: use summ_by_year_state
# aggregated to the state level

#-- Dependent vs Explanatory --#

# means by state
color_value <- "Mean Median Inc ($T)"

ggplot(summ_by_year_state, aes(y=wmean_mh_days, x=wmean_inequality, color=wmean_medinc/1000)) + 
  geom_point() +
  geom_smooth(formula = y ~ x, method=lm, se=FALSE, color="#a89968") + 
  labs(title = "Scatterplot of Population-weighted Means by State",
       y = 'Poor Mental Health Days per 30', x = 'Income Inequality Rate', 
       color = color_value)

ggplot(summ_by_year_state, aes(y=wmean_mh_rate, x=wmean_inequality, color=wmean_medinc/1000)) + 
  geom_point() +
  geom_smooth(formula = y ~ x, method=lm, se=FALSE, color="#a89968") + 
  labs(title = "Scatterplot of Population-weighted Means by State",
       y = 'Mental Distress Rate', x = 'Income Inequality Rate', 
       color = color_value)

# medians by state
color_value <- "Med. Median Inc ($T)"

ggplot(summ_by_year_state, aes(y=med_mh_days, x=med_inequality, color=med_medinc/1000)) + 
  geom_point() +
  geom_smooth(formula = y ~ x, method=lm, se=FALSE, color="#a89968") +  ## Add linear regression line
  labs(title = "Scatterplot of Medians by State",
       y = 'Poor Mental Health Days per 30', x = 'Income Inequality Rate', 
       color = color_value)

ggplot(summ_by_year_state, aes(y=med_mh_rate, x=med_inequality, color=med_medinc/1000)) + 
  geom_point() +
  geom_smooth(formula = y ~ x, method=lm, se=FALSE, color="#a89968") + 
  labs(title = "Scatterplot of Medians by State",
       y = 'Mental Distress Rate', x = 'Income Inequality Rate', 
       color = color_value)


```

[INSERT brief summary of histograms]

```{r hist_multiple, results='markup', eval=F}

# https://www.r-bloggers.com/2016/07/quick-plot-of-all-variables/

keep_var <- c('inequality', 'unempl', 'child_poverty', 'hs_grad', 'college', 'single_parent', 'severe_housing', 'food_index', 'mental_health_days', 'mental_distress_rate', 'median_inc')

#ranked[keep_var]

ranked[keep_var] %>% 
  gather() %>% 
  ggplot(aes(x=value)) + 
    facet_wrap(~ key, scales = "free") + 
    geom_histogram(bins=50)


```

```{r histograms, results='markup', eval=F}

# Shumel -- please work on some histograms


```

[INSERT brief summary of boxplots]

```{r boxplots, results='markup'}

##Assigned to: Xuan
# In this part, in order to better show the relationship between the economic level and mental health of people, mental_distress_rate in different regions of the United States, I divided the United States into 4 regions by geographical method (for details, you can check this link below). I will analyze the four parts of "Northeast", "South", "West", "Midwest" separately.(https://www2.census.gov/geo/pdfs/maps-data/maps/reference/us_regdiv.pdf) In addition, I would like to say that there are many independent variables and dependent variables selected by our group, but in order to shorten the space and highlight the key points, I only select two independent variables and two dependent variable for analysis. The two independent variables that I selected are "Median household income" (median_inc) and "Income inequality" (inequality). The two dependent variable that I selected is "Poor mental health days" (mental_health_days) and "Frequent mental distress rate" (mental_distress_rate).
dframe0NA <- na.omit(dframe)
# In the above code, I remove all NA of the dframe and save it as dframe0NA.
#View(dframe0NA)
dframe2016 = subset(dframe0NA, year == 2016)
dframe2017 = subset(dframe0NA, year == 2017)
dframe2018 = subset(dframe0NA, year == 2018)
dframe2019 = subset(dframe0NA, year == 2019)
dframe2020 = subset(dframe0NA, year == 2020)
dframe2021 = subset(dframe0NA, year == 2021)
# In the above code, I divided dframe0NA into 6 subsets by year, from 2016 to 2021. This makes subsequent drawings easier.
dframeNortheast = subset(dframe0NA, region == "Northeast")
dframeSouth = subset(dframe0NA, region == "South")
dframeWest = subset(dframe0NA, region == "West")
dframeMidwest = subset(dframe0NA, region == "Midwest")
# In the above code, I divided dframe0NA into 4 subsets by regions. This makes subsequent drawings easier.

loadPkg("ggplot2")
dframe$year <- factor(dframe$year)
# In the above code, I set "year" as a variable. Frankly speaking, this step can be done earlier. In the below codes, because what I went to draw was some boxplots, which usually only represents one variable, I designed the code in order to be able to relate the variable to different regions (the U.S. is divided into four geographic regions). In each box plot, you can see the one variable in different regions.
```

```{r boxplots_median_inc, results='markup'}
##Assigned to: Xuan
##### Characterizations of the Median household income of four regions in the United States from 2016 to 2021.
##### The following 6 photographs, show the median household income of 4 regions from 2016 to 2021 (6 years). So I mean, one photograph is related to one year. And the trend of these 6 photographs are same. It is easy to find that the median household income of northeast region is the biggest. Followed by west region. It's not difficult to understand, the northeast region has New York, the biggest city in US, the west region has LA, the second biggest city in US. In addition, the median household income of midwest region and south region are relatively low. But the gap between rich and poor in the midwest and south is relatively small.

# Characterization of the Median household income of four regions in the United States in 2016.
h1 <- ggplot(dframe2016, aes(x=region, y=median_inc)) + 
  geom_boxplot() + 
  geom_boxplot(colour="orange",fill="#7777cc",outlier.colour="red",outlier.shape=8, outlier.size=4) +
  labs(title="2016")

# Characterization of the Median household income of four regions in the United States in 2017.
h2 <-ggplot(dframe2017, aes(x=region, y=median_inc)) + 
  geom_boxplot() + 
  geom_boxplot(colour="orange",fill="#7777cc",outlier.colour="red",outlier.shape=8, outlier.size=4) +
  labs(title="2017")

# Characterization of the Median household income of four regions in the United States in 2018.
h3 <-ggplot(dframe2018, aes(x=region, y=median_inc)) + 
  geom_boxplot() + 
  geom_boxplot(colour="orange",fill="#7777cc",outlier.colour="red",outlier.shape=8, outlier.size=4) +
  labs(title="2018")

# Characterization of the Median household income of four regions in the United States in 2019.
h4 <-ggplot(dframe2019, aes(x=region, y=median_inc)) + 
  geom_boxplot() + 
  geom_boxplot(colour="orange",fill="#7777cc",outlier.colour="red",outlier.shape=8, outlier.size=4) +
  labs(title="2019")

# Characterization of the Median household income of four regions in the United States in 2020.
h5 <-ggplot(dframe2020, aes(x=region, y=median_inc)) + 
  geom_boxplot() + 
  geom_boxplot(colour="orange",fill="#7777cc",outlier.colour="red",outlier.shape=8, outlier.size=4) +
  labs(title="2020")

# Characterization of the Median household income of four regions in the United States in 2021.
h6 <-ggplot(dframe2021, aes(x=region, y=median_inc)) + 
  geom_boxplot() + 
  geom_boxplot(colour="orange",fill="#7777cc",outlier.colour="red",outlier.shape=8, outlier.size=4) +
  labs(title="2021")

histplot1 <- grid.arrange(h1,h2,h3,h4,h5,h6, nrow=2, ncol=3)
annotate_figure(histplot1, top = text_grob("The Characterization of Median Income by Region per Year", color = "black", face = "bold", size = 14))
```

```{r boxplots_median_inc_region, results='markup'}
##### Characterizations of the Median household income of four regions from 2016 to 2021, decided by regions.
##### From the four photographs below, we can see that the trend of median household income in all 4 regions gradually upward from 2016 to 2021. For example, in south region, the median household income in 2016 was about $40,000, and by 2021, it was close to $50,000.

# Characterization of the Median household income of northeast region from 2016 to 2021.
ggplot(dframeNortheast, aes(x=year, y=median_inc)) +
  geom_boxplot() +
  geom_boxplot(colour="orange",fill="#7777cc",outlier.colour="red",outlier.shape=8, outlier.size=4) +
  labs(title="The trend of median_inc of northeast region from 2016 to 2021")

# Characterization of the Median household income of south region from 2016 to 2021.
ggplot(dframeSouth, aes(x=year, y=median_inc)) +
  geom_boxplot() +
  geom_boxplot(colour="orange",fill="#7777cc",outlier.colour="red",outlier.shape=8, outlier.size=4) +
  labs(title="The trend of median_inc of south region from 2016 to 2021")

# Characterization of the Median household income of west region from 2016 to 2021.
ggplot(dframeWest, aes(x=year, y=median_inc)) +
  geom_boxplot() +
  geom_boxplot(colour="orange",fill="#7777cc",outlier.colour="red",outlier.shape=8, outlier.size=4) +
  labs(title="The trend of median_inc of west region from 2016 to 2021")

# Characterization of the Median household income of midwest region from 2016 to 2021.
ggplot(dframeMidwest, aes(x=year, y=median_inc)) +
  geom_boxplot() +
  geom_boxplot(colour="orange",fill="#7777cc",outlier.colour="red",outlier.shape=8, outlier.size=4) +
  labs(title="The trend of median_inc of midwest region from 2016 to 2021")
```

```{r boxplots_inequality, results='markup'}
##Assigned to: Xuan
##### Characterizations of Income inequality of four regions in the United States from 2016 to 2021.
##### The following 6 photographs, show the income inequality of 4 regions from 2016 to 2021 (6 years). So I mean, one photograph is related to one year. And the trend of these 6 photographs are same. From these 6 photographs, we can find that the income inequality of south region is biggest.This also happens to validate the six plots above for median household income. The South, which has the lowest median household income, is likely to have the greatest income inequality problem.

# Characterization of Income inequality of four regions in the United States in 2016.
ggplot(dframe2016, aes(x=region, y=inequality)) + 
  geom_boxplot() + 
  geom_boxplot(colour="orange",fill="#7777cc",outlier.colour="red",outlier.shape=8, outlier.size=4) +
  labs(title="The Characterization of Income inequality of four regions in 2016")

# Characterization of Income inequality of four regions in the United States in 2017.
ggplot(dframe2017, aes(x=region, y=inequality)) + 
  geom_boxplot() + 
  geom_boxplot(colour="orange",fill="#7777cc",outlier.colour="red",outlier.shape=8, outlier.size=4) +
  labs(title="The Characterization of Income inequality of four regions in 2017")

# Characterization of Income inequality of four regions in the United States in 2018.
ggplot(dframe2018, aes(x=region, y=inequality)) + 
  geom_boxplot() + 
  geom_boxplot(colour="orange",fill="#7777cc",outlier.colour="red",outlier.shape=8, outlier.size=4) +
  labs(title="The Characterization of Income inequality of four regions in 2018")

# Characterization of Income inequality of four regions in the United States in 2019.
ggplot(dframe2019, aes(x=region, y=inequality)) + 
  geom_boxplot() + 
  geom_boxplot(colour="orange",fill="#7777cc",outlier.colour="red",outlier.shape=8, outlier.size=4) +
  labs(title="The Characterization of Income inequality of four regions in 2019")

# Characterization of Income inequality in four regions of the United States in 2020.
ggplot(dframe2020, aes(x=region, y=inequality)) + 
  geom_boxplot() + 
  geom_boxplot(colour="orange",fill="#7777cc",outlier.colour="red",outlier.shape=8, outlier.size=4) +
  labs(title="The Characterization of Income inequality of four regions in 2020")

# Characterization of Income inequality of four regions in the United States in 2021.
ggplot(dframe2021, aes(x=region, y=inequality)) + 
  geom_boxplot() + 
  geom_boxplot(colour="orange",fill="#7777cc",outlier.colour="red",outlier.shape=8, outlier.size=4) +
  labs(title="The Characterization of Income inequality of four regions in 2021")
```

```{r boxplots_inequality_region, results='markup'}
##### Characterizations of the income inequality of four regions from 2016 to 2021, decided by regions.
##### From the four photographs below, we can easily find that, no matter which region it is, the status of income inequality has not changed much from 2016 to 2021. For example, in the Northeast region, it has been around 4.4% for 6 consecutive years. This kind of data result could provide some advice for local governments to continue working to reduce or even eliminate income inequality.

# Characterization of the income inequality of northeast region from 2016 to 2021.
ggplot(dframeNortheast, aes(x=year, y=inequality)) +
  geom_boxplot() +
  geom_boxplot(colour="orange",fill="#7777cc",outlier.colour="red",outlier.shape=8, outlier.size=4) +
  labs(title="The trend of Income inequality of northeast region from 2016 to 2021")

# Characterization of the income inequality of south region from 2016 to 2021.
ggplot(dframeSouth, aes(x=year, y=inequality)) +
  geom_boxplot() +
  geom_boxplot(colour="orange",fill="#7777cc",outlier.colour="red",outlier.shape=8, outlier.size=4) +
  labs(title="The trend of Income inequality of south region from 2016 to 2021")

# Characterization of the income inequality of west region from 2016 to 2021.
ggplot(dframeWest, aes(x=year, y=inequality)) +
  geom_boxplot() +
  geom_boxplot(colour="orange",fill="#7777cc",outlier.colour="red",outlier.shape=8, outlier.size=4) +
  labs(title="The trend of Income inequality of west region from 2016 to 2021")

# Characterization of the income inequality of midwest region from 2016 to 2021.
ggplot(dframeMidwest, aes(x=year, y=inequality)) +
  geom_boxplot() +
  geom_boxplot(colour="orange",fill="#7777cc",outlier.colour="red",outlier.shape=8, outlier.size=4) +
  labs(title="The trend of Income inequality of midwest region from 2016 to 2021")
```

```{r boxplots_mh_days, results='markup'}
##Assigned to: Xuan
##### Characterizations of the Poor mental health days/month of four regions in the United States from 2016 to 2021.
##### In the following 6 photographs, we can easily find that the south region has the biggest poor mental health days. This is consistent with our previous findings. The South has the lowest median household income of the four regions. So, people in the South are more likely to be unhappy.

# Characterization of the Poor mental health days/month of four regions in the United States in 2016.
h7<-ggplot(dframe2016, aes(x=region, y=mental_health_days)) + 
  geom_boxplot() + 
  geom_boxplot(colour="orange",fill="#7777cc",outlier.colour="red",outlier.shape=8, outlier.size=4) +
  labs(title="2016")

# Characterization of the Poor mental health days/month of four regions in the United States in 2017.
h8<-ggplot(dframe2017, aes(x=region, y=mental_health_days)) + 
  geom_boxplot() + 
  geom_boxplot(colour="orange",fill="#7777cc",outlier.colour="red",outlier.shape=8, outlier.size=4) +
  labs(title="2017")

# Characterization of the Poor mental health days/month of four regions in the United States in 2018.
h9<-ggplot(dframe2018, aes(x=region, y=mental_health_days)) + 
  geom_boxplot() + 
  geom_boxplot(colour="orange",fill="#7777cc",outlier.colour="red",outlier.shape=8, outlier.size=4) +
  labs(title="2018")

# Characterization of the Poor mental health days/month of four regions in the United States in 2019.
h10<-ggplot(dframe2019, aes(x=region, y=mental_health_days)) + 
  geom_boxplot() + 
  geom_boxplot(colour="orange",fill="#7777cc",outlier.colour="red",outlier.shape=8, outlier.size=4) +
  labs(title="2019")

# Characterization of the Poor mental health days/month of four regions in the United States in 2020.
h11<-ggplot(dframe2020, aes(x=region, y=mental_health_days)) + 
  geom_boxplot() + 
  geom_boxplot(colour="orange",fill="#7777cc",outlier.colour="red",outlier.shape=8, outlier.size=4) +
  labs(title="2020")

# Characterization of the Poor mental health days/month of four regions in the United States in 2021.
h12<-ggplot(dframe2021, aes(x=region, y=mental_health_days)) + 
  geom_boxplot() + 
  geom_boxplot(colour="orange",fill="#7777cc",outlier.colour="red",outlier.shape=8, outlier.size=4) +
  labs(title="2021")

histplot2 <- grid.arrange(h7,h8,h9,h10,h11,h12, nrow=2, ncol=3)
annotate_figure(histplot2, top = text_grob("The Characterization of Mentally Unhealthy Days/Month by Region per Year", color = "black", face = "bold", size = 14))

```

```{r boxplots_mh_days_region, results='markup'}
##Assigned to: Xuan
##### Characterizations of the Poor mental health days/month of 4 regions from 2016 to 2021
##### From following 4 photographs, we can easily find that the poor mental health days gradually go up, no matter in which region. Combined with the results of the previous findings, even though the income of local people is increasing year by year, their unhappy time is still increasing. My guess is that income growth is not keeping up with local prices.

# Characterization of the Poor mental health days/month of northeast region from 2016 to 2021.
ggplot(dframeNortheast, aes(x=year, y=mental_health_days)) + 
  geom_boxplot() + 
  geom_boxplot(colour="orange",fill="#7777cc",outlier.colour="red",outlier.shape=8, outlier.size=4) +
  labs(title="mental unhealthy days/month of northeast region from 2016 to 2021")

# Characterization of the Poor mental health days/month of south region from 2016 to 2021.
ggplot(dframeSouth, aes(x=year, y=mental_health_days)) + 
  geom_boxplot() + 
  geom_boxplot(colour="orange",fill="#7777cc",outlier.colour="red",outlier.shape=8, outlier.size=4) +
  labs(title="mental unhealthy days/month of south region from 2016 to 2021")

# Characterization of the Poor mental health days/month of west region from 2016 to 2021.
ggplot(dframeWest, aes(x=year, y=mental_health_days)) + 
  geom_boxplot() + 
  geom_boxplot(colour="orange",fill="#7777cc",outlier.colour="red",outlier.shape=8, outlier.size=4) +
  labs(title="mental unhealthy days/month of west region from 2016 to 2021")

# Characterization of the Poor mental health days/month of midwest region from 2016 to 2021.
ggplot(dframeMidwest, aes(x=year, y=mental_health_days)) + 
  geom_boxplot() + 
  geom_boxplot(colour="orange",fill="#7777cc",outlier.colour="red",outlier.shape=8, outlier.size=4) +
  labs(title="mental unhealthy days/month of midwest region from 2016 to 2021")
```

```{r boxplots_md_rate, results='markup'}
##Assigned to: Xuan
##### Characterizations of the Frequent mental distress rate of four regions in the United States from 2016 to 2021.
##### The following 6 photographs shows the similar results as the above "Characterizations of the Poor mental health days/month of four regions in the United States from 2016 to 2021". Stress, nervousness, and mentally unhealthy emotions are all related to income. The lower the income, the more stressful it is. Just like south region.

# Characterization of the Frequent mental distress rate of four regions in the United States in 2016.
h13<-ggplot(dframe2016, aes(x=region, y=mental_distress_rate)) + 
  geom_boxplot() + 
  geom_boxplot(colour="orange",fill="#7777cc",outlier.colour="red",outlier.shape=8, outlier.size=4) +
  labs(title="2016")

# Characterization of the Frequent mental distress rate of four regions in the United States in 2017.
h14<-ggplot(dframe2017, aes(x=region, y=mental_distress_rate)) + 
  geom_boxplot() + 
  geom_boxplot(colour="orange",fill="#7777cc",outlier.colour="red",outlier.shape=8, outlier.size=4) +
  labs(title="2017")

# Characterization of the Frequent mental distress rate of four regions in the United States in 2018.
h15<-ggplot(dframe2018, aes(x=region, y=mental_distress_rate)) + 
  geom_boxplot() + 
  geom_boxplot(colour="orange",fill="#7777cc",outlier.colour="red",outlier.shape=8, outlier.size=4) +
  labs(title="2018")

# Characterization of the Frequent mental distress rate of four regions in the United States in 2019.
h16<-ggplot(dframe2019, aes(x=region, y=mental_distress_rate)) + 
  geom_boxplot() + 
  geom_boxplot(colour="orange",fill="#7777cc",outlier.colour="red",outlier.shape=8, outlier.size=4) +
  labs(title="2019")

# Characterization of the Frequent mental distress rate of four regions in the United States in 2020.
h17<-ggplot(dframe2020, aes(x=region, y=mental_distress_rate)) + 
  geom_boxplot() + 
  geom_boxplot(colour="orange",fill="#7777cc",outlier.colour="red",outlier.shape=8, outlier.size=4) +
  labs(title="2020")

# Characterization of the Frequent mental distress rate of four regions in the United States in 2021.
h18<-ggplot(dframe2021, aes(x=region, y=mental_distress_rate)) + 
  geom_boxplot() + 
  geom_boxplot(colour="orange",fill="#7777cc",outlier.colour="red",outlier.shape=8, outlier.size=4) +
  labs(title="2021")

histplot3 <- grid.arrange(h13,h14,h15,h15,h17,h18, nrow=2, ncol=3)
annotate_figure(histplot3, top = text_grob("The Characterization of Frequent Mental Distress Rate by Region per Year", color = "black", face = "bold", size = 14))

```

```{r boxplots_md_rate_region, results='markup'}
##Assigned to: Xuan
##### Characterizations of the Frequent mental distress rate of 4 regions from 2016 to 2021
##### From following 4 photographs, we can easily find that the frequent mental distress rate gradually goes up, no matter in which region. Combined with the results of the previous findings, even though the income of local people is increasing year by year, it is still easy for then to feel stressed. My guess is that income growth is not keeping up with local prices. Maybe because of COVID-19, we can see that the trend from 2020 to 2021 is more obvious than previous years.

  geom_boxplot() + 
  geom_boxplot(colour="orange",fill="#7777cc",outlier.colour="red",outlier.shape=8, outlier.size=4) +
  labs(title="Frequent mental distress rate of northeast region from 2016 to 2021")

  geom_boxplot() + 
  geom_boxplot(colour="orange",fill="#7777cc",outlier.colour="red",outlier.shape=8, outlier.size=4) +
  labs(title="Frequent mental distress rate of south region from 2016 to 2021")

  geom_boxplot() + 
  geom_boxplot(colour="orange",fill="#7777cc",outlier.colour="red",outlier.shape=8, outlier.size=4) +
  labs(title="Frequent mental distress rate of west region from 2016 to 2021")

  geom_boxplot() + 
  geom_boxplot(colour="orange",fill="#7777cc",outlier.colour="red",outlier.shape=8, outlier.size=4) +
  labs(title="Frequent mental distress rate of midwest region from 2016 to 2021")
```

## Measures of Variation
*When applicable.*

**Assigned to:** Xuan

Variance, standard deviation, etc.

```{r variance, results='markup'}

# Variance, standard of median_inc of different region from 2016 to 2021.
var(dframeNortheast$median_inc)
var(dframeSouth$median_inc)
var(dframeWest$median_inc)
var(dframeMidwest$median_inc)

sd(dframeNortheast$median_inc)
sd(dframeSouth$median_inc)
sd(dframeWest$median_inc)
sd(dframeMidwest$median_inc)
#From above data, we can see the Northeast region's people has the largest gap between the rich and the poor, and the Midwest has the smallest gap.

# Variance, standard of inequality of different region from 2016 to 2021.
var(dframeNortheast$inequality)
var(dframeSouth$inequality)
var(dframeWest$inequality)
var(dframeMidwest$inequality)

sd(dframeNortheast$inequality)
sd(dframeSouth$inequality)
sd(dframeWest$inequality)
sd(dframeMidwest$inequality)
#From above data, we can see the South region's people has the largest income inequality.

# Variance, standard of mental_distress_rate of different region from 2016 to 2021.
var(dframeNortheast$mental_distress_rate)
var(dframeSouth$mental_distress_rate)
var(dframeWest$mental_distress_rate)
var(dframeMidwest$mental_distress_rate)

sd(dframeNortheast$mental_distress_rate)
sd(dframeSouth$mental_distress_rate)
sd(dframeWest$mental_distress_rate)
sd(dframeMidwest$mental_distress_rate)
#From above data, we can see the South region's people has the largest mental distress (maybe because of low income), and the Northeast has the smallest mental distress (maybe because they have much more money).

# Variance, standard of mental_health_days of different region from 2016 to 2021.
var(dframeNortheast$mental_health_days)
var(dframeSouth$mental_health_days)
var(dframeWest$mental_health_days)
var(dframeMidwest$mental_health_days)

sd(dframeNortheast$mental_health_days)
sd(dframeSouth$mental_health_days)
sd(dframeWest$mental_health_days)
sd(dframeMidwest$mental_health_days)
#From above data, we can see the South region's people has the largest mental unhealthy days/month (maybe because of low income), and the Northeast has the smallest mental unhealthy days/month (maybe because they have much more money).

```


## Normality tests
*When applicable.*

**Assigned to:** Shumel

For example, histograms, Q-Q plots, etc. Try `ezids::outlierKD2()` function if necessary.

```{r norm, results='markup', eval=F}

qqnorm()
qqline()

```



## Initial Correlation and Hypothesis Tests
*When applicable.*

**Assigned to:** Alex (with help from Mark)

For example, Chi Square tests, ANOVA analysis, Z-test or Z-interval, T-test or T-interval, etc.

```{r corr, results='markup', eval=F}

t.test()

table() ## creates contingency table
chisq.test()

aov(y ~ x, data=)

sapply(dframe2, class)

a<-pairs(dframe2))
b<-cor(a) ## correlation matrix
corrplot(dframe2)


```


```{r estimation, results='markup'}

# Estimation: Establish confidence intervals
# t-interval by region for inequality, mental_health_days, mental_distress_rate

# estimate for all regions
ttest99_in <- t.test(x=ranked$inequality, conf.level=0.99)
ttest99_mh_days <- t.test(x=ranked$mental_health_days, conf.level=0.99)
ttest99_mh_rate <- t.test(x=ranked$mental_distress_rate, conf.level=0.99)

ci99_in <- paste0("[", paste0(round(ttest99_in$conf.int, 4), collapse = ", "), "]")

ci99_mh_days <- paste0("[", paste0(round(ttest99_mh_days$conf.int, 4), collapse = ", "), "]")

ci99_mh_rate <- paste0("[", paste0(round(ttest99_mh_rate$conf.int, 4), collapse = ", "), "]")

var_list <- c("inequality", "mental_health_days", "mental_distress_rate")

ci_list <- list(rep(c("All"), times = length(var_list)), var_list, c(ci99_in, ci99_mh_days, ci99_mh_rate))


# estimate by region
region_list <- c("Midwest", "Northeast", "South", "West")

for (r in region_list) {  ## region names
  
  ranked_region <- ranked %>% subset(region==r)
  
  ttest99_in <- t.test(ranked_region$inequality, conf.level=0.99)
  ttest99_mh_days <- t.test(ranked_region$mental_health_days, conf.level=0.99)
  ttest99_mh_rate <- t.test(ranked_region$mental_distress_rate, conf.level=0.99)
  
  ci99_in <- paste0("[", paste0(round(ttest99_in$conf.int, 4), collapse = ", "), "]")
  
  ci99_mh_days <- paste0("[", paste0(round(ttest99_mh_days$conf.int, 4), collapse = ", "), "]")
  
  ci99_mh_rate <- paste0("[", paste0(round(ttest99_mh_rate$conf.int, 4), collapse = ", "), "]")
  
  ci_list[[1]] <- append(ci_list[[1]], rep(c(r), times = length(var_list)))
  ci_list[[2]] <- append(ci_list[[2]], var_list)
  ci_list[[3]] <- append(ci_list[[3]], c(ci99_in, ci99_mh_days, ci99_mh_rate))
  
  #print(ci_list)
}

# print results
for (n in 1:15) {
  
  print(paste(ci_list[[1]][n], ci_list[[2]][n], ci_list[[3]][n], sep=" | "))
  
}

```

Estimate the t-intervals of our key variables at the 99% confidence level.

Given the large size of our data set, 

mean estimates


The t-intervals by region are reported for each variable of interest in the following table:

***
Region | Variable | 99% conf.int |
-------|----------|--------------|
`r ci_list[[1]][1]` | `r ci_list[[2]][1]` | `r ci_list[[3]][1]` |
`r ci_list[[1]][2]` | `r ci_list[[2]][2]` | `r ci_list[[3]][2]` |
`r ci_list[[1]][3]` | `r ci_list[[2]][3]` | `r ci_list[[3]][3]` |
`r ci_list[[1]][4]` | `r ci_list[[2]][4]` | `r ci_list[[3]][4]` |
`r ci_list[[1]][5]` | `r ci_list[[2]][5]` | `r ci_list[[3]][5]` |
`r ci_list[[1]][6]` | `r ci_list[[2]][6]` | `r ci_list[[3]][6]` |
`r ci_list[[1]][7]` | `r ci_list[[2]][7]` | `r ci_list[[3]][7]` |
`r ci_list[[1]][8]` | `r ci_list[[2]][8]` | `r ci_list[[3]][8]` |
`r ci_list[[1]][9]` | `r ci_list[[2]][9]` | `r ci_list[[3]][9]` |
`r ci_list[[1]][10]` | `r ci_list[[2]][10]` | `r ci_list[[3]][10]` |
`r ci_list[[1]][11]` | `r ci_list[[2]][11]` | `r ci_list[[3]][11]` |
`r ci_list[[1]][12]` | `r ci_list[[2]][12]` | `r ci_list[[3]][12]` |
`r ci_list[[1]][13]` | `r ci_list[[2]][13]` | `r ci_list[[3]][13]` |
`r ci_list[[1]][14]` | `r ci_list[[2]][14]` | `r ci_list[[3]][14]` |
`r ci_list[[1]][15]` | `r ci_list[[2]][15]` | `r ci_list[[3]][15]` |
***


```{r}

# Hypothesis Testing: Test the process of making decisions about the value of a population parameter

# difference in means
# - rural vs urban (create binary indicator at 50%?)
# - across report years (2017 to 2021)
# - by region (ANOVA or chi-sq??)



```




## Linear Modeling and Regression
*When applicable. Might not need at all.*

**Assigned to:** TBD

```{r reg, results='markup', eval=F}

lm()
xkablevif()
anova()
xkabledply()


```

