---
title: "Technical Analysis"
author: "Team 6"
#date: "3/23/2022"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    number_sections: false
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
---

```{r init, include=F}

# Import libraries
library(ezids)
#library(readr)
library(dplyr)
library(data.table)

```


```{r setup, include=FALSE}

# knitr::opts_chunk$set(warning = F, results = "markup", message = F)
knitr::opts_chunk$set(warning = F, results = "hide", message = F, echo = TRUE)
options(scientific=T, digits = 3) 
# options(scipen=9, digits = 3) 
# ‘scipen’: integer. A penalty to be applied when deciding to print numeric values in fixed or exponential notation.  Positive values bias towards fixed and negative towards scientific notation: fixed notation will be preferred unless it is more than ‘scipen’ digits wider.
# use scipen=999 to prevent scientific notation at all times

```


# Technical Analysis

## Instructions

Technical analysis (R code), 25% (First attempt due: Mar 28 Monday 12:30pm)

Just before your team presentation, provide an R-markdown, or R script file, which shows the R-code and brief explanations as well as the rationale of the Exploratory Data Analysis of
your project. (Also submit your data file, or give the online source url.) This submission,
together with your final document (in part VI), will determine your team’s grade. This
document shows a technical person the math/stat/codes that you used in your analysis. It
should include:  
- Summary of the dataset
- Descriptive Statistics
- Graphical representations of the data
- [When applicable] Measures of Variance / sd
- [When applicable] Normality tests
- [When applicable] Initial correlation / Chi Square tests / ANOVA analysis / Z-test or Z-interval / T-test or T-interval etc.


## Summary

**Assigned to:** Mark

```{r import, results='hide', echo=F}

# import csv as dataframe
dframe <- data.frame(read.csv("data/processed/analytic_data_2016_2021.csv"))

# convert variables to type factor
dframe$statecode <- factor(dframe$statecode)
dframe$countycode <- factor(dframe$countycode)
dframe$fipscode <- factor(dframe$fipscode)
dframe$county_ranked <- factor(dframe$county_ranked)

# check structure of data
str(dframe)

```

```{r summary, results='hide'}

# look at county_ranked var; not all counties are ranked; also some aggregated data per state and country exist in the observations
# =1 means they are ranked, =0 means unranked, and =NA is for state/national data
print(summary(dframe$county_ranked))

# subset of dataframe including only ranked counties
ranked <- dframe %>% subset(county_ranked==1)


xkabledplyhead(ranked)



```

Our data set contains `r nrow(dframe)` observations and `r ncol(dframe)` variables.



## Descriptive Statistics

**Assigned to:** TBD

```{r dstats, results='markup'}



#summary()
#xkablesummary()


```


## Graphs and Figures

**Assigned to:** TBD

```{r graphs, results='markup'}


```


## Measures of Variation
*When applicable.*

**Assigned to:** TBD

Variance, standard deviation, etc.

```{r variance, results='markup'}


```


## Normality tests
*When applicable.*

**Assigned to:** TBD

For example, 

```{r norm, results='markup'}


```



## Initial Correlation and Hypothesis Tests
*When applicable.*

**Assigned to:** TBD

For example, Chi Square tests, ANOVA analysis, Z-test or Z-interval, T-test or T-interval, etc.

```{r corr, results='markup'}


```



## Linear Modeling and Regression
*When applicable. Might not need at all.*

**Assigned to:** TBD

```{r reg, results='markup'}


```

